---
title: "Seurat_Part4_dataextraction"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## R Markdown

```{r}
source("https://bioconductor.org/biocLite.R")
#biocLite("BiocUpgrade")
biocLite("digest")
#biocLite("ggplot2")
#biocLite("DESeq2")
#biocLite("zoo")
#biocLite("Seurat")
library(Seurat)
dir<-"Volumes/Eythora/backup/singlecelljuly2018/Pilot2018"
setwd(dir)
library(ggplot2)
filename<-"yourfilename_Part3.RData"
load(file=filename)
experiment.aggregate
```

Build a cluster tree, and look at the relative location of each cluster. 
```{r}
experiment.aggregate <- SetAllIdent(experiment.aggregate, id = "finalcluster")
use.pcs=20
experiment.aggregate<- BuildClusterTree(
  experiment.aggregate,
  pcs.use = use.pcs,
  do.reorder = F,
  reorder.numeric = F,
  do.plot=F)
PlotClusterTree(experiment.aggregate, cex=1, use.edge.length=F,node.depth=2)
table(experiment.aggregate@ident)
table(experiment.aggregate@meta.data$MouseID,experiment.aggregate@meta.data$finalcluster)
```
Based on the tree above, you can combine neighboring clusters. 
```{r}
Rename_Multiple<-function(object, oldnames, newnames,index=1:length(oldnames)){
        for (i in index){
                object<-RenameIdent(
                        object=object,
                        old.ident.name = oldnames[i],
                        new.ident.name = newnames[i])
        }
        object
}
experiment.aggregate <- SetAllIdent(experiment.aggregate, id = "finalclusters")
oldcluster<-c("microglia-8")
newcluster<-c("microglia-4")
experiment.aggregate<-Rename_Multiple(experiment.aggregate,oldcluster,newcluster)
experiment.aggregate <- AddMetaData(
  object = experiment.aggregate,
  metadata = experiment.aggregate@ident, #make a new metadata column containing the final set of clusters as shown in @ident
  col.name = "finalcluster")
table(experiment.aggregate@meta.data$finalcluster)
```
Find markers for the new clusters (with names)

```{r}
markers_all <- FindAllMarkers(
    object = experiment.aggregate, 
    only.pos = TRUE, 
    min.pct = 0.25, #gene must be present in 25% of the cells in the cluster
    logfc.threshold = 0.25
)
dim(markers_all)
head(markers_all)
write.csv(markers_all,"RNUMIallmarkerssortonllowfilternobatch.csv")
table(table(markers_all$gene)) #basically a histogram of how many genes are present in how many clusters

markers_all_single <- markers_all[markers_all$gene %in% names(table(markers_all$gene))[table(markers_all$gene) == 1],] #markers that define a single cluster
write.csv(markers_all_single,"RNUMIsinglemarkerssortonllowfilternobatch.csv")
dim(markers_all_single)
table(table(markers_all_single$gene))

table(markers_all_single$cluster)

head(markers_all_single)
```
Get means of each marker gene per cluster (in the cluster, out of the cluster)
```{r}
getGeneClusterMeans <- function(gene, cluster){
  x <- experiment.aggregate@data[gene,]
  m <- tapply(x, ifelse(experiment.aggregate@ident == cluster, 1, 0), mean)
  mean.in.cluster <- m[2]
  mean.out.of.cluster <- m[1]
  return(list(mean.in.cluster = mean.in.cluster, mean.out.of.cluster = mean.out.of.cluster))
}

## for sake of time only using first six (head)
means <- mapply(getGeneClusterMeans, markers_all[,"gene"], markers_all[,"cluster"])
means <- matrix(unlist(means), ncol = 2, byrow = T)

colnames(means) <- c("mean.in.cluster", "mean.out.of.cluster")
markers_all2 <- cbind(markers_all, means)
head(markers_all2)
write.csv(markers_all2,"markers_all_numiregress_names_nobatch.csv")
```



Adding in a new metadata column representing samples within clusters

```{r}
samplecluster = paste(experiment.aggregate@meta.data$orig.ident,experiment.aggregate@meta.data$finalcluster,sep = '-')
mouseIDfull = paste(experiment.aggregate@meta.data$orig.ident, experiment.aggregate@meta.data$MouseID)

names(mouseIDfull) = rownames(experiment.aggregate@meta.data)
names(samplecluster) = rownames(experiment.aggregate@meta.data)

head(samplecluster) # tells you the cells in each sample in each cluster

experiment.aggregate <- AddMetaData( 
  object = experiment.aggregate,
  metadata = samplecluster,
  col.name = "samplecluster")

experiment.aggregate <- AddMetaData( 
  object = experiment.aggregate,
  metadata = mouseIDfull,
  col.name = "mouseIDfull")
 
```

Subsetting samples - could extract a particular cluster using your new metadata column
```{r}
subset="yourchosencelltype"
experiment.subset<- SubsetData(
  object = experiment.aggregate,
  subset.raw=T,
  do.clean=T, #do.clean removes things like scaledata, tsne, pca so that you can recalculate them below
  cells.use = rownames(experiment.aggregate@meta.data)[experiment.aggregate@meta.data$celltype %in% subset])

experiment.subset <- FindVariableGenes(
  object = experiment.subset,
  mean.function = ExpMean,
  dispersion.function = LogVMR,
  x.low.cutoff = 0.125,
  x.high.cutoff = 5,
  y.cutoff = 0.25, do.plot=T) #1468 genes

length(experiment.subset@var.genes)

experiment.subset <- ScaleData(
  object = experiment.subset,
  genes.use=experiment.subset@var.genes)

experiment.subset <- RunPCA(
  object = experiment.subset,
  pc.genes = experiment.subset@var.genes,
  do.print = TRUE,
  pcs.print = 1:5,
  genes.print = 5,
  pcs.compute = 40,
  maxit = 500)

PrintPCAParams(experiment.subset)
  
PCElbowPlot(
  experiment.subset,
  num.pc = 40)
  
use.pcs = 1:20 #set to whatever you chose from the previous section

#will take a long time to run
experiment.subset <- FindClusters(
    object = experiment.subset, 
    reduction.type = "pca", 
    dims.use = use.pcs, 
    resolution = seq(0.5,4,0.5), #pick multiple resolutions; smaller = fewer clusters seq(0.5,4,0.5), #
    print.output = FALSE, 
    save.SNN = TRUE,
    force.recalc = TRUE
)
PrintFindClustersParams(object = experiment.subset)


sapply(grep("^res",colnames(experiment.subset@meta.data),value = TRUE),
       function(x) length(unique(experiment.subset@meta.data[,x])))

experiment.subset <- SetAllIdent(experiment.subset, id = "res.1.5") #assign a cluster ID to each cell; to color with higher resolution, change this id to the resolution you care about

table<-table(experiment.subset@ident,experiment.subset@meta.data$orig.ident)
depvctrl<-table[,1]/table[,2]
table<-cbind(table,depvctrl)
table
total<-table(experiment.subset@meta.data$orig.ident)
ratio<-total[1]/total[2]
names(ratio)<-"expected"
ratio<-rep(ratio,length(depvctrl))
clusterbias<-as.data.frame(cbind(table,ratio))

bias<-function(table){
  bias<-rep("no",nrow(table))
  for (i in 1:nrow(table)){
    bias[i]="Both"
    if (table[i,3]>2*table[i,4]){
      bias[i]="Dep"
    }
    else if (table[i,3]<table[i,4]/2){
      bias[i]="Ctrl"
    }
  }
  table<-cbind(table,bias)
}

#assign cluster identities based on ratio
clusterbias<-bias(clusterbias)
clusterbias
clusterbias[,c(3,5)]
clusterid<-experiment.subset@ident
levels(clusterid)<-clusterbias$bias
cellbias<-clusterid
experiment.subset <- AddMetaData(
        object = experiment.subset,
        metadata = cellbias,
        col.name= "bias")
table(experiment.subset@meta.data$bias,experiment.subset@meta.data$orig.ident)

experiment.subset <- RunTSNE(
  object = experiment.subset,
  reduction.use = "pca",
  dims.use = use.pcs,
  seed.use=10)
TSNEPlot(object = experiment.subset, pt.size=0.5,do.label=T)
TSNEPlot(object = experiment.subset, group.by="orig.ident", pt.size=0.5)
experiment.subset<- BuildClusterTree(
  experiment.subset,
  pcs.use = use.pcs,
  do.reorder = F,
  reorder.numeric = F,
  do.plot=T)
#edit names as needed
oldnames<-c("0","1","2","3","4","5","6")
newnames<-c("Dep-1","Both-1","Both-2","Dep-2","Ctrl-1","Ctrl-2","Dep-3")
PlotClusterTree(experiment.subset, cex=1, use.edge.length=F,node.depth=2)
Rename_Multiple<-function(object, oldnames, newnames,index=1:length(oldnames)){
        for (i in index){
                object<-RenameIdent(
                        object=object,
                        old.ident.name = oldnames[i],
                        new.ident.name = newnames[i])
        }
        object
}
experiment.subset<-Rename_Multiple(experiment.subset,oldnames,newnames)
save(experiment.subset,file="subset.RData")
```



#rownames(experiment.aggregate@meta.data) are cells
#TSNEPlot(object = experiment.sort_contra, group.by="ident", pt.size=0.5, do.label = F)

#FeaturePlot(experiment.sort_contra, features.plot=c('Sparcl1'), pt.size=0.5)
#FeaturePlot(experiment.sort_contra, features.plot=c('Sparc'), pt.size=0.5)



```{r}

table(experiment.subset@meta.data$bias,experiment.subset@meta.data$celltype)

#TSNEPlot(object = experiment.sort_ipsi, group.by="ident", pt.size=0.5, do.label = F)

#FeaturePlot(experiment.sort_ipsi, features.plot=c('Sparcl1'), pt.size=0.5)
#FeaturePlot(experiment.sort_ipsi, features.plot=c('Sparc'), pt.size=0.5)
```


```{r}
filename<-"yournamehere_Part4.RData"
save(list=ls(),file=filename)
```

```{r}
sessionInfo()
```

